BEX3012 Project Report \n Detecting Facial Expressions in Professional Tennis Matches

### 2017-06-18 Internship/BEX3012ProjectReport.Rmd
Presentation Changes

---
title: "BEX3012 Project Report \n Detecting Facial Expressions in Professional Tennis Matches"
author: "Stephanie Kobakian"
date: "28 March 2017"
output: bookdown::pdf_document2
fig_caption: yes
out_width: 7
fontsize: 11pt
link-citations: true
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)

library(tidyverse)
library(readr)
library(GGally)
library(magick)
library(grid)
library(gridExtra)
library(ggplot2)
library(stargazer)
library(glmnet)

#Read in data files
ME <- read_csv("data/MicrosoftEmotions.csv")
SB <- read_csv("data/SkybiometryEmotions.csv")
GE <- read_csv("data/GoogleFacesEmotions.csv")
GF <- read_csv("data/GoogleFaces.csv")
DE <- read_csv("data/DominantEmotion.csv")

@@

```{r mutate}
#Create normalised GE values
Conf <- function(Emot){
ifelse(Emot=="VERY_UNLIKELY", 10,  
        ifelse(Emot == "UNLIKELY", 30, 
                ifelse(Emot == "POSSIBLE", 50, 
                ifelse(Emot == "LIKELY", 70, 
                ifelse(Emot == "VERY_LIKELY", 90, NA)))))
}

@@

```{r join}
ME <- left_join(ME, GF, by = "FileName") 
SB <- left_join(SB, GF, by = c("aname"="FileName")) 
```

```{r playerSubset}
GEP <- GE %>% dplyr::filter(detect == "Player")
MEP <- ME %>% dplyr::filter(detect == "Player")
SBP <- SB %>% dplyr::filter(detect == "Player")
```

```{r predominantEmotion}
#shows only the rank 1 emotion and value for each face
GEP1 <- GEP %>% 
  filter(!is.na(anger)) %>%
  group_by(FileName) %>% 
  gather(key = Emotion, value=EmotionValue, 
         anger, happiness, sadness, surprise, neutral)  %>% 
  arrange(FileName, desc(EmotionValue)) %>%
      group_by_(~ FileName) %>% 
      slice(1)

@@

```{r NeutralProportions}
S4 <- select(SBP, FileName = aname, Skybiometry = neutral)
S5 <- select(MEP, FileName, Microsoft = neutral)
S6 <- select(GEP, FileName, Google = neutral)
J3 <- full_join(S4, S5, by = "FileName")
J4 <- full_join(J3, S6, by = "FileName")
```

```{r regressionFunctions}

@@

\newpage
# Introduction

Facial Recognition is beginning to be explored in sports environments, this presents quality issues.

helpful for applications

@@

# Materials

@@

The image set used to derive emotion information contained 6406 still shot images. These were 800x450px video frames, taken every 3 seconds from 5 minute video segments. These segments were sampled from 105 Australian Open 2016 match broadcast videos.
The broadcast videos are a combination of multiple camera feeds, these angles vary depending on the court the match was played on.

```{r ExampleImage, echo=F, fig.cap = "This image was derived from a match played between JMillman and DSchwartzmann in the first Round of the 2016 Australian Open. \\label{fig:ExampleImage}"}

img <- image_read("Figures/2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0020.png")

@@

Three emotion recognition APIS were considered.
They were chosen for their accessibility and online reviews of their performance, and these are well recognised options for Facial recognition currently available.

```{r Emotion-Software-Specs-Table, echo=F, fig.cap = "\\label{tab:Emotion-Software-Specs-Table}"}

knitr::kable(EmotionSoftwareSpecs, longtable=TRUE, booktabs = TRUE, caption="This details the capabilities we considered important in recognising emotions in images of faces.")
```

As can be seen above, a noticeable difference between the three is the amount of times the API can be called within a given time frame. Skybiometry had the largest imposition on Call Limits as it only allowed 100 API calls to be processed per hour. Microsoft also had a limit imposed, but this allowed for much more to be processed with the possibility of 1200 images to be processed within one hour, after accounting for the wait time between each group of 20.
Google Vision's API call limit had a minimal effect for our purposes.

The number of emotions shown in the results were not congruent. These differences made it necessary to process the outputs to provide the most comparable results for analysis.
Microsoft provided the most emotion options, contempt being the addition that was not provided by Skybiometry. Google provided the least, only providing four different emotions.

The observations of these emotions from the APIs were also different. Google provides likelihoods of an emotion occurring on a particular face in categorical options ranging from Very Unlikely to Very Likely. Microsoft provide Proportions, ranging between 0 and 1, whereas Skybiometry results in a Confidence value for the emotion occurring in the specified face in values from 0 to 100.

\newpage
# Procedure

@@

To create a set of individual faces for the APIs to consider we derived small images that were cropped subsets of the 450x800px images.
The crop was performed to create a new image with a unique name that focused on the area marked as a face by Google^[based on Google’s bounding box coordinates]. A border to frame the face was also included for visual appeal.
These 1319 new images, of varying sizes, were hosted on Google Drive as individual images.

```{r VisualSet, echo=F, fig.cap = "Three Faces in the image set, after extracting faces from the full broadcast video stills, and the dominant emotion each software recognised for them. \\label{fig:VisualSet}"}

@@

  dl <- d %>% 
  select(Skybiometry = SEmotion, Microsoft = MEmotion, Google = GEmotion) %>%
    gather(key = Software, value = Emotion, Microsoft, Skybiometry, Google)

  t <- tableGrob(dl, theme = ttheme_minimal(core=list(fg_params=list(hjust=0, x=0.1)),
                      rowhead=list(fg_params=list(hjust=0, x=0))), rows = c("", "", ""))

@@

grid.arrange(i1, i2, i3, t1, t2, t3, nrow=2)

@@

As mentioned previously, the results extracted for each face differed across the APIs this made it necessary to produce a comparable set of results for analysis.

Microsoft provided a POST request object in a JSON format. 
This was easily manipulated to create a data set of the eight numeric emotion values provided for each face. In the event that emotion values were not returned for a certain face, a row of NAs will be associated with the unique face ID in the data set.

```{r MEexample, echo=F, message=F, fig.cap = "\\label{tab:MEexample}"}

MEexample <- ME %>% subset(FileName=="face-1745-1-Go.png")

knitr::kable(MEexample[2:9], digits=6, longtable=TRUE, booktabs = TRUE, caption="The Microsoft API provided eight numerical values, one for each emotion.")
```

Skybiometry also provided a POST request object in a JSON format. However, unlike Microsoft, the results included a True or False Value for each of the seven emotions and a 'confidence value as a percentage from 0 to 100'^[https://skybiometry.com/documentation/#document-21] representing the confidence of the emotion being present on a player's face.

```{r SBexample, echo=F, fig.cap = "\\label{tab:SBexample}"}

SBexample <- SB %>% subset(aname=="face-1745-1-Go.png") %>% select(ends_with(".confidence"))

knitr::kable(SBexample[3:9], digits=6, longtable=TRUE, booktabs = TRUE,
             col.names = c("neutral", "anger", "disgust", "fear",
                           "happiness", "sadness", "surprise"),
             caption="The Skybiometry API provided seven numerical Confidence values, one for each emotion.")
```

@@

```{r GEexample, echo=F, fig.cap = "\\label{tab:GEexample}"}
GEexample <- GE %>% subset(FileName=="face-1745-1-Go.png") %>% select(ends_with("Likelihood"))

@@

Noticing that the sums of the individual emotions for each face in the Microsoft results was approximately one provided the basis for transformation for the other APIs. Dividing the individual Skybiometry emotion confidences by the sum of the confidences would give the proportional likelihood of each emotion being the dominant emotion in a face.

@@

```{r SBT, echo=F, fig.cap = "\\label{tab:SBT}"}
SBT <- SB %>% subset(aname=="face-1745-1-Go.png") %>% select(neutral, anger, disgust, fear, happiness, sadness, surprise)

knitr::kable(SBT, digits=6, longtable=TRUE, booktabs = TRUE,
               caption="The Skybiometry API provided 7 numerical Confidence values, one for each emotion.")
```

@@

knitr::kable(LT, longtable=TRUE, booktabs = TRUE, caption="")
```

@@

The process to derive the numeric values for Google followed the process undertaken for Skybiometry values. Where the individual emotion confidences were divided by the sum of the emotion confidence values for each face. This resulted in congruent emotion data.

@@

GET <- GEP %>% subset(FileName=="face-1745-1-Go.png") %>% 
  select(happiness, sadness, anger, surprise, neutral)

knitr::kable(GET, longtable=TRUE, booktabs = TRUE,
             col.names = c("happiness", "sadness", "anger", "surprise", "neutral"),
             caption="The Google Vision API provided four likelihoods, one for each emotion.")
```

\newpage
# Analysis

Each API returned a different amount of non null results. Microsoft produced emotion results for `r sum(!is.na(ME$anger))` faces. This was more than Skybiometry, as it produced only `r sum(!is.na(SB$anger))` results for emotions on faces. Google provided the most, with  `r sum(!is.na(GE$anger))` emotion results^[As the initial set was faces identified by Google there is the possibility of bias toward being able to identify emotions in faces it was able to find originally.].

```{r playerNonplayer, eval=FALSE,fig.cap="\\label{tab:playerNonPlayer}"}
#Apply a machine learning algorithm to face recognition data, and software provided for fitting model, aiming to classify whether an identified face is a player or not player
ME <- ME %>% rowwise %>% mutate(Player = ifelse(detect=="Player", "Player","NotPlayer")) %>%  mutate(Player01 = ifelse(Player=="Player",1,0))

@@

gpm <- model.matrix(gplayer)

gpmNet = glmnet(gpm, as.vector(GE$Player01))

pot(gpmNet)

#cv of model
cvgpmNet = cv.glmnet(gpm, as.vector(GE$Player01),family = "binomial", type.measure = "class")

plot(cvgpmNet)
summary(gplayer)
coef(cvgpmNet, s = "lambda.min")
```

```{r NonNullFaces, fig.cap = "\\label{fig:NonNullFaces}", results = 'asis'}

J4 <-J4 %>% rowwise() %>%
  mutate(Gna = ifelse(!is.na(Google), TRUE, FALSE)) %>%
  mutate(Sna = ifelse(!is.na(Skybiometry), TRUE, FALSE)) %>%
  mutate(Mna = ifelse(!is.na(Microsoft), TRUE, FALSE))

stargazer(format(nfft, quote=FALSE, justify="right"),title = "Emotion results were found by all three APIs for 364 common faces, none of the APIs produced results for 10 player faces. Google was able to produce 382 results that Skybiometry and Microsoft did not.", header=FALSE, type="latex")
```

In the following sections we will only consider the faces of players, as they would be the intended targets for our applications of this research. Microsoft produced emotion results for `r sum(!is.na(MEP$anger))` player's faces. Skybiometry, produced `r sum(!is.na(SBP$anger))` and Google provided `r sum(!is.na(GEP$anger))` emotion results.

@@

```{r GEPhists, fig.cap = "It can be seen that there are significant proportions of faces where the non neutral emotions occur close to zero. This is balanced by the large percentage, over 75%, where the neutral value of just under .75 occurred. The faces used showed higher levels of happiness than any other non neutral emotion as the percentage occuring at zero was less than the percentages of the other emotions.\\label{fig:GEPhists}"}
gganG <- ggplot(GEP, aes(x=anger)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggjoG <- ggplot(GEP, aes(x=happiness)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggneG <- ggplot(GEP, aes(x=neutral)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggsoG <- ggplot(GEP, aes(x=sadness)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggsuG <- ggplot(GEP, aes(x=surprise)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

grid.arrange(gganG,ggjoG,ggneG,ggsoG,ggsuG,nrow=2,
             top=textGrob("Google Distribution of Emotion Porportions",
             gp=gpar(fontsize=20,font=3)), 
             left = ("Percentage of Emotion values"))

@@

```{r SBPhists, fig.cap = "The distributions show that over 70% of faces had levels of happiness and neutrality at, or close to, 0. The distributions show that there are very little faces with specific emotion levels above 50%. Instead there are low levels of all emotions found on many faces. \\label{fig:SBPhists}"}
gganS <- ggplot(SBP, aes(x=anger))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

gghaS <- ggplot(SBP, aes(x=happiness))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggneS <- ggplot(SBP, aes(x=neutral))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggsaS <- ggplot(SBP, aes(x=sadness))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggsuS <- ggplot(SBP, aes(x=surprise))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggdiS <- ggplot(SBP, aes(x=disgust))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

ggfeS <- ggplot(SBP, aes(x=fear))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

grid.arrange(gganS, gghaS, ggneS, ggsaS, ggsuS, ggdiS, ggfeS,nrow=3,
             top=textGrob("Skybiometry Distribution of Emotion Porportions",gp=gpar(fontsize=20,font=3)), 
             left = ("Percentage of Emotion values"))
```

```{r MEPhists, fig.cap="There were many faces that returned high neutral values. This is matched by the high percentages of faces where the proportion levels were at, or close to, zero. While the happiness emotion was most varied, with values between 0 and 1, over 40% of faces returned surprise levels above zero.\\label{fig:MEPhists}"}

gganM <- ggplot(MEP, aes(x=anger)) +  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) + coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

gghaM <- ggplot(MEP, aes(x=happiness)) +  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) + coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")

>>>>

---
title: "BEX3012 Project Report \n Detecting Facial Expressions in Professional Tennis Matches"
author: "Stephanie Kobakian"
date: "12 May 2017"
output: bookdown::pdf_document2
fig_caption: yes
out_width: 7
fontsize: 10pt
link-citations: true
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache=FALSE)

library(tidyverse)
library(readr)
library(GGally)
library(magick)
library(grid)
library(gridExtra)
library(ggplot2)
library(glmnet)
library(stargazer)

#Read in data files
ME <- read_csv("data/MicrosoftEmotions.csv")
SB <- read_csv("data/SkybiometryEmotions.csv")
GE <- read_csv("data/GoogleFacesEmotions.csv")
GF <- read_csv("data/GoogleFaces.csv")
DE <- read_csv("data/DominantEmotions.csv")
DEne<-DE

@@

```{r mutate, cache =FALSE}
#Create normalised GE values
Conf <- function(Emot){
ifelse(Emot=="VERY_UNLIKELY", 10,  
        ifelse(Emot == "UNLIKELY", 30, 
                ifelse(Emot == "POSSIBLE", 50, 
                ifelse(Emot == "LIKELY", 70, 
                ifelse(Emot == "VERY_LIKELY", 90, NA)))))
}

@@

```{r join, cache=FALSE}
ME <- left_join(ME, GF, by = "FileName") 
SB <- left_join(SB, GF, by = c("aname"="FileName")) 
```

```{r playerSubset, cache=FALSE}
GEP <- GE %>% dplyr::filter(detect == "Player")
MEP <- ME %>% dplyr::filter(detect == "Player")
SBP <- SB %>% dplyr::filter(detect == "Player")
```

```{r predominantEmotion, cache=FALSE}
#shows only the rank 1 emotion and value for each face
GEP1 <- GEP %>% 
  filter(!is.na(anger)) %>%
  group_by(FileName) %>% 
  gather(key = Emotion, value=EmotionValue, 
         anger, happiness, sadness, surprise, neutral)  %>% 
  arrange(FileName, desc(EmotionValue)) %>%
      group_by_(~ FileName) %>% 
      slice(1)

@@

```{r NeutralProportions, cache = FALSE}
S4 <- select(SBP, FileName = aname, Skybiometry = neutral)
S5 <- select(MEP, FileName, Microsoft = neutral)
S6 <- select(GEP, FileName, Google = neutral)
J3 <- full_join(S4, S5, by = "FileName")
J4 <- full_join(J3, S6, by = "FileName")
```

```{r regressionFunctions, cache =FALSE}

@@

@@

Tennis is often considered a mental game due to the pressures placed on the two, or four, individuals during match play. This report explores the application of emotion recognition technology to elite tennis match broadcasts. This extends on research of the facial detection performance of these APIs. By understanding the emotions expressed on a face we have the opportunity to consider a player's mental state as a match progresses. This would help players to improve their performance if correlations can be found between match performance and their emotional state.
By analysing a players mental state from observations rather than self reporting we may be able to gain information that a player will not recall after a match. It is predicted that the software will be able to derive the emotional expression shown by a player. It is expected that emotional recognition may be impaired by certain features of the face captured.
This research exposes the impact of presenting APIs with faces captured from unexpected angles and in unusual poses, and how this may be impacting the performance of the APIs. The analyses undertaken show the differing results returned from the three APIs used, validation of these results was not achieveable within the time frame of this project.

\newpage
# Materials

@@

The image set used to derive emotion information contained 6406 still shot images. These were 800x450px video frames, taken every 3 seconds from 5 minute video segments. These segments were sampled from 105 Australian Open 2016 match broadcast videos.
The broadcast videos are the televised match play shown on Channel Seven. They contain shots from various cameras, these angles vary depending on the court the match was played on.

```{r ExampleImage, echo=F, fig.cap = "This image of John Millman was taken from the broadcast video of a match played against DSchwartzmann in the first Round of the 2016 Australian Open. This image is representative of the most commmon group of faces found. It captures a player's face close up on the court, taken at the player's shoulder height, with the logo wall in the back ground. \\label{fig:ExampleImage}"}

img <- image_read("Figures/2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0013.png")

@@

Three emotion recognition APIs were considered.
They were chosen for their accessibility and online reviews of their performance. Each API is a well recognised Facial and Emotion recognition solution currently available online.

```{r Emotion-Software-Specs-Table, echo=F,results = 'asis', fig.cap = "\\label{tab:Emotion-Software-Specs-Table}"}

knitr::kable(EmotionSoftwareSpecs, longtable=TRUE, booktabs = TRUE, caption="This table details the capabilities that were considered in contrasting the accessibility and outputs of APIs that recognise emotions in images of faces.")
```

The three APIs reviewed in this report have been summarised in Table \@ref(tab:Emotion-Software-Specs-Table) above. A noticeable difference between the three is the amount of times the API can be called within a given time frame. Skybiometry had the largest imposition on Call Limits as it only allowed 100 API calls to b e processed per hour. Microsoft also had a limit imposed, but this allowed for much more to be processed with the possibility of 1200 images to be processed within one hour, after accounting for the wait time between each group of 20.
Google Vision's API call limit had a minimal effect for our purposes.

There were a different set of emotions in the results provided by each API.
These differences made it necessary to process the outputs to provide the most comparable results for analysis.
Microsoft provided the most emotional categories, 'contempt' being the addition that was not provided by Skybiometry or Google. Google provided the least, only providing four emotions, joy, sorrow, surprise and anger.

The observations of these emotions from the APIs were also different. Google provides likelihoods of an emotion occurring on a particular face in categorical options ranging from Very Unlikely to Very Likely. Microsoft provide Proportions, ranging between 0 and 1, the sum of the emotion values for each face summed to approximately one. Whereas Skybiometry results in a confidence value of the emotion occurring in the specified face in values from 0 to 100, these values were not mutually exclusive and resulted in a different sum across each face.

Cost and Access refers to the steps that must be undertaken for a new user to begin Emotion Recognition using each API. All options require an account be created with a current email address. Google and Skybiometry require payments to be made. 
The Google Vision API provides Facial and Emotion Recognition for $1.50 for Units 1001 - 5,000,000 per month^[https://cloud.Google.com/vision/docs/pricing].
Skybiometry have three API subscription options that vary in the amount of API calls that can be made per month, day and hour^[https://skybiometry.com/pricing/].

# Procedure

@@

To create a set of individual faces for the APIs to consider we derived small images that were subsets of the 450x800px images.
The crop was performed to focus on the area marked as a face by Google^[based on Google’s bounding box coordinates].
These 1319 new images, of varying sizes, were hosted on Google Drive as individual images.

```{r VisualSet, echo=F, fig.cap = "This figure provides three faces in the image set, after extracting faces from the full broadcast video stills. Below the faces are the APIs predominant emotion category recognised in the face. \\label{fig:VisualSet}"}

@@

  dl <- d %>% 
  select(Skybiometry = Skybiometry, Microsoft = Microsoft, Google = Google) %>%
    gather(key = Software, value = Emotion, Microsoft, Skybiometry, Google)

  t <- tableGrob(dl, theme = ttheme_minimal(core=list(fg_params=list(hjust=0, x=0.1)),rowhead=list(fg_params=list(hjust=0, x=0))), rows = c("", "", ""))

@@

grid.arrange(i1, i2, i3, t1, t2, t3, nrow=2)
```

@@

For each individual face a POST request as submitted to each API, these returned POST request objects in a JSON format. When this object contained values for the emotional state of a face we considered this to be a 'response'. Where the alternative was NA results being returned.

Given the different forms of the outputs produced by each API, we felt a  harmonization process was required prior to undertaking a comparative analysis. The goal of the harmonization was to produce Emotion values for each face that would be comparable across the three APIs.

Microsoft provided a POST request object in a JSON format. 
This was easily manipulated to create a data set of the eight numeric emotion values provided for each face. In the event that emotion values were not returned for a certain face, a row of NAs is associated with the unique face ID in the data set. This data set contained a column for the following emotions: anger, contempt, disgust, fear, happiness,
neutral, sadness, surprise.

```{r MEexample, echo=F, results = 'asis', message=F, fig.cap = "\\label{tab:MEexample}"}

MEexample <- ME %>% subset(FileName=="face-1635-1-Go.png")

MEexample <- lapply(MEexample[2:9], function(x){replace(x, x <0.001,0)})

knitr::kable(as.data.frame(MEexample),longtable=TRUE, booktabs = TRUE, digits=3, format.args = list(nsmall=3, digits=3, scientific=FALSE), caption="The Microsoft API provided eight numerical values, one for each emotion.")
```

Table \@ref(tab:MEexample) details the emotion category levels for the first face in Figure \@ref(fig:VisualSet). The highest emotion level is for surprise, and the lowest value of 2.78e-05 is associated with contempt.

Skybiometry, unlike Microsoft, the results included a True or False Value for each of the seven emotions and a 'confidence value as a percentage from 0 to 100'^[https://skybiometry.com/documentation/#document-21] representing the confidence of the emotion being present on a player's face. Table \@ref(tab:SBexample) contains the confidence values for the first face in Figure \@ref(fig:VisualSet). If a confidence value was 50 or above, the emotion 'value' would be labelled True. This would have occurred for surprise and disgust.

```{r SBexample, echo=F, results = 'asis', fig.cap = "\\label{tab:SBexample}"}

SBexample <- SB %>% subset(aname=="face-1635-1-Go.png") %>% select(ends_with(".confidence"))

knitr::kable(SBexample[3:9], longtable=TRUE, booktabs = TRUE,col.names = c("neutral", "anger", "disgust", "fear", "happiness", "sadness", "surprise"),
             caption="The Skybiometry API provided seven numerical Confidence values, one for each emotion.")
```

@@

```{r GEexample, echo=F, fig.cap = "\\label{tab:GEexample}"}
GEexample <- GE %>% subset(FileName=="face-1635-1-Go.png") %>% select(ends_with("Likelihood"))

@@

The same face that had produced high levels of surprise for Microsoft and Skybiometry produced the results in Table \@ref(tab:GEexample). The most positive result was the 'Unlikely' categorisation for surprise, this proposes that surprise was more prominent than other emotions but less significant than the other APIs intimated.

\newpage
### API Result Transformations

Noticing that the sums of the individual emotions for each face in the Microsoft results was approximately one, provided the basis for transformation for the other APIs. Dividing the individual Skybiometry emotion confidences by the sum of the confidences would give the proportional likelihood of each emotion being the dominant emotion in a face.

@@

```{r SBT, echo=F, fig.cap = "\\label{tab:SBT}"}
SBT <- SB %>% subset(aname=="face-1635-1-Go.png") %>% select(neutral, anger, disgust, fear, happiness, sadness, surprise)

SBTexample <- lapply(SBT, function(x){replace(x, x <0.001,0)})

knitr::kable(as.data.frame(SBTexample), longtable=TRUE, booktabs = TRUE, digits=3, caption="The Skybiometry API provided 7 numerical Confidence values, one for each emotion.")
```

@@

knitr::kable(LT,longtable=TRUE, booktabs = TRUE, caption="")
```

@@

The process to derive the numeric values for Google followed the steps undertaken for Skybiometry values. Where the individual emotion confidences were divided by the sum of the emotion confidence values for each face. This transformation results in the values seen in Table \@ref(tab:GET), the faces can now be analysed numerical and compared to the other APIs numerical results.

@@

GET <- GEP %>% subset(FileName=="face-1635-1-Go.png") %>% 
  select(happiness, sadness, anger, surprise, neutral)

GETexample <- lapply(GET, function(x){replace(x, x <0.001,0)})

knitr::kable(as.data.frame(GETexample), longtable=TRUE, booktabs = TRUE, digits=3, caption="The Google Vision API provided five proportions, one for each emotion and a neutral category.")
```

### Predominant Emotion Categorisations

To simplify the analysis, we assigned a predominant emotion for each API result. This allowed for clearer comparisons of the emotion results for individual faces across the three APIs.
For Google, Skybiometry and Microsoft, the emotion proportion levels for each individual face were considered, the maximum proportion value of each APIs emotion set was deemed the dominant emotion for each face. Recalling that there is a difference in the amount of emotion categories considered by each software.

For each individual face this resulted in three dominant emotion classifications, one per API.
The dominant emotion would be one of seven emotions for Google, one of eight for Microsoft, and one of five for Google.

\newpage
# Results

Each API returned a different amount of total responses. Microsoft produced emotion results for `r sum(!is.na(ME$anger))` faces, 55.497% of the faces provided.
This was more than Skybiometry, as it produced only `r sum(!is.na(SB$anger))` results for emotions on faces. This is 48.103% of the set of faces.
Google provided the most, 94.162% of the faces, `r sum(!is.na(GE$anger))`, were given emotion results^[As the initial set was faces identified by Google there is the possibility of bias toward being able to identify emotions in faces it was able to find originally.].

```{r NonNullFaces, fig.cap = "\\label{fig:NonNullFaces}", results = 'asis'}
J4 <- J4 %>% rowwise() %>%
  mutate(Gna = ifelse(!is.na(Google), TRUE, FALSE)) %>%
  mutate(Sna = ifelse(!is.na(Skybiometry), TRUE, FALSE)) %>%
  mutate(Mna = ifelse(!is.na(Microsoft), TRUE, FALSE))

#Gneu <- DE %>% filter(is.na(Microsoft) & is.na(Skybiometry))
#table(Gneu$Google)

stargazer(format(nfft, quote=FALSE, justify="right"),title = "Emotion results were found by all three APIs for 364 common faces, none of the APIs produced results for 10 player faces. Google was able to produce 380 results that Skybiometry and Microsoft did not.", header=FALSE, type="latex")
```

### Player Faces

```{r playerNonplayer,results = 'asis', fig.cap=" \\label{tab:playerNonPlayer}"}
#Apply a machine learning algorithm to face recognition data, and software provided for fitting model, aiming to classify whether an identified face is a player or not player
ME <- ME %>% rowwise %>% mutate(Player = ifelse(detect=="Player", "Player","NotPlayer")) %>%  mutate(Player01 = ifelse(Player=="Player",1,0)) 
GE <- GE %>% rowwise %>% mutate(Player = ifelse(detect=="Player", "Player","NotPlayer")) %>%  mutate(Player01 = ifelse(Player=="Player",1,0))

@@

gplconfint <- as.data.frame(confint(gplayer)) %>% 
  mutate(`exp(2.5%)` = exp(`2.5 %`)) %>% 
  mutate(`exp(97.5%)` = exp(`97.5 %`)) %>%
  cbind(rownames(confint(gplayer)), exp(gplayer$coefficients), .) %>%
  select(`exp(Estimates)` = `exp(gplayer$coefficients)`, `exp(2.5%)`, `exp(97.5%)`)

knitr::kable(gplconfint[2:9,] ,longtable=TRUE, booktabs = TRUE, digits=3, caption="A logisitic regression was undertaken to show how certain image attributes contribute to the likelihood of the face being a player.")
```

The regression coefficients are the change in the likelihood of a face being a player if the category is true for a face. The most significant predictors of a response are outlined in the table above. It can be seen that when the image was taken at a player's "Shoulder Height" then it is, on average, 4.45 times more likely to be a "Player" than a fan or staff member on court.
When the head angle is "Other", it is between 1.34 and 3.87 times more likely to be a Player than faces that have been captured front on, at the 5% level of confidence.

In the following sections we will only consider the faces of players, as they would be the intended targets for our applications of this research. Microsoft produced emotion results for `r sum(!is.na(MEP$anger))` player's faces. Skybiometry, produced `r sum(!is.na(SBP$anger))` and Google provided `r sum(!is.na(GEP$anger))` emotion results for faces.

@@

```{r GEPhists, fig.width=9, fig.cap = "It can be seen that there are significant proportions of faces where the non neutral emotions occur close to zero. This is balanced by the large percentage, over 75%, where the neutral value of just under .75 occurred. The faces used showed higher levels of happiness than any other non neutral emotion as the percentage occuring at zero was less than the percentages of the other emotions.\\label{fig:GEPhists}"}
gganG <- ggplot(GEP, aes(x=anger)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("") + theme(axis.text.x = element_text(angle = 20))

gghaG <- ggplot(GEP, aes(x=happiness)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("") + theme(axis.text.x = element_text(angle = 20))

ggneG <- ggplot(GEP, aes(x=neutral)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("") + theme(axis.text.x = element_text(angle = 20))

ggsaG <- ggplot(GEP, aes(x=sadness)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("") + theme(axis.text.x = element_text(angle = 20))

ggsuG <- ggplot(GEP, aes(x=surprise)) + 
  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("") + theme(axis.text.x = element_text(angle = 20))

grid.arrange(ggneG,
             gganG,
             gghaG,
             ggsaG,
             ggsuG,
             nrow=2,
             top=textGrob("Google Distribution of Emotion Porportions",
             gp=gpar(fontsize=20,font=3)), 
             left = ("Percentage of Emotion values"))

@@

```{r SBPhists, fig.width=9, fig.cap = "The distributions show that over 70% of faces had levels of happiness and neutrality at, or close to, 0. The distributions show that there are very little faces with specific emotion levels above 50%. Instead there are low levels of all emotions found on many faces. \\label{fig:SBPhists}"}
gganS <- ggplot(SBP, aes(x=anger))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("") + theme(axis.text.x = element_text(angle = 20))

gghaS <- ggplot(SBP, aes(x=happiness))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")+ theme(axis.text.x = element_text(angle = 20))

ggneS <- ggplot(SBP, aes(x=neutral))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")+ theme(axis.text.x = element_text(angle = 20))

ggsaS <- ggplot(SBP, aes(x=sadness))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")+ theme(axis.text.x = element_text(angle = 20))

ggsuS <- ggplot(SBP, aes(x=surprise))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")+ theme(axis.text.x = element_text(angle = 20))

ggdiS <- ggplot(SBP, aes(x=disgust))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")+ theme(axis.text.x = element_text(angle = 20))

ggfeS <- ggplot(SBP, aes(x=fear))+  geom_histogram(binwidth=0.055, aes(y=..count../sum(..count..))) +coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + ylab("")+ theme(axis.text.x = element_text(angle = 20))

grid.arrange(ggneS,
             gganS, 
             ggdiS, 
             ggfeS, 
             gghaS, 
             ggsaS, 
             ggsuS,nrow=2,
             top=textGrob("Skybiometry Distribution of Emotion Porportions",gp=gpar(fontsize=20,font=3)), 
             left = ("Percentage of Emotion values"))
```

```{r MEPhists,fig.width=9, fig.cap="There were many faces that returned high neutral values. This is matched by the high percentages of faces where the proportion levels were at, or close to, zero. While the happiness emotion was most varied, with values between 0 and 1, over 40% of faces returned surprise levels above zero.\\label{fig:MEPhists}"}
